Visualization and Unsupervised Classification of Water Composition with UAV-based Hyperspectral Imaging and Generative Topographic Mapping

### 2024-05-17 paper/paper.tex

A Freefly Alta-X autonomous quadcopter was used as the UAV platform in this study. This UAV was equipped with a Resonon Pika XC2 visible+near-infrared (VNIR) hyperspectral imager to capture HSI with $462$ wavelengths per pixel ranging from $391$ to $1011$ nm. This imager is in a pushbroom configuration so that HSI are captured one scan-line at a time resulting in data cubes consisting of $1000$ scan-lines with $1600$ pixels each. Additionally, the camera includes an embedded GPS/INS unit to enable georectification of collected HSI. An upward facing Ocean Optics UV-Vis NIR spectrometer with a cosine corrector was also included to provide measurements of the incident solar irradiance spectrum. Data collection and processing was controlled by an attached Intel NUC small-form-factor computer.

@@

@@

%\begin{figure}[t]
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{paper/figures/results/bic.pdf}
%\end{adjustwidth}
\caption{Results of the hyperparameter search. Left: Variation of BIC with $m$ and $s$ for fixed $\alpha=0.1$. Right: Variation of the BIC with with $m$ and $\alpha$ for fixed $s=1.0$. The white star in each plot indicates the parameters with the lowest BIC across the entire parameter search.\label{fig:hp-results}}
\end{figure}

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{paper/figures/results/spectral-signatures.png}
\end{adjustwidth}
\caption{Spectral signatures $\psi(\xi_k)$ corresponding to GTM nodes with non-zero responsibility for exemplar spectra corresponding to the rhodamine dye plume (top left), algae (top right), dry grass (bottom left), and open water (bottom right). A pseudo-color image is inset into each plot wit the location of the exemplar spectrum marked with a white circle. \label{fig:spectral-signatures}}
\end{figure}

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{paper/figures/results/algae.png}
\end{adjustwidth}
\caption{Using the spectral endmember assigned to algae from the trained GTM together with the NS3, we are able to estimate the algal abundance near the shore. (\textbf{left}) NS3 values of pixels below a threshold of 0.4275 corresponding to a total area of $587.3$ $m^2$. (\textbf{right}) A picture of the pond near the shore showing the algae.\label{fig:algae-map}}
\end{figure}

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{paper/figures/results/rhodamine.png}
\end{adjustwidth}
\caption{Using the spectral endmember assigned to Rhodamine from the trained GTM together with the NS3, we are able to map the evolution of the dye plume across two UAV flights. (\textbf{left}) The initial dye plume corresponding to a total area of $330.7$ $m^2$. (\textbf{right}) The same dye plume imaged approximately 15 minutes later now subsumes a total area of $1495.6$ $m^2$.\label{fig:rhodamine-map}}
\end{figure}

@@

\noindent 
\begin{tabular}{@{}ll}
UAV & Unmanned Aerial Vehicle \\
SOM & Self Organizing Map \\ 
GTM & Generative Topographic Mapping  \\
HSI & Hyperspectral Image \\ 
CDOM & Colored Dissolved Organic Matter \\
HAB & Harmful Algal Bloom 
\end{tabular}
}

>>>>

A Freefly Alta-X autonomous quadcopter was used as the UAV platform in this study. This UAV was equipped with a Resonon Pika XC2 visible+near-infrared (VNIR) hyperspectral imager to capture HSI with $462$ wavelengths per pixel ranging from $391$ to $1011$ nm. This imager is in a pushbroom configuration so that HSI are captured one scan-line at a time resulting in data cubes consisting of $1000$ scan-lines with $1600$ pixels each. Additionally, the camera includes an embedded GPS/INS unit to enable georectification of collected HSI. An upward facing Ocean Optics UV-Vis NIR spectrometer with a cosine corrector was also included to provide measurements of the incident solar irradiance spectrum. The configuration of the UAV with the attached hyperspectral imager is shown in Fig~\ref{fig:drone}. Data collection and processing was controlled by an attached Intel NUC small-form-factor computer.

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{paper/figures/methods/annotated-drone.pdf}
\end{adjustwidth}
\caption{The UAV platform: (\textbf{a}) The Resonon Pika XC2 hyperspectral imager. (\text{b}) The Freefly Alta-X with the attached hyperspectral imager, processing computer, and downwelling irradiance spectrometer. \label{fig:drone}}
\end{figure}

@@

%\begin{figure}[t]
\begin{figure}[h]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=17.0cm]{paper/figures/results/bic.pdf}
\end{adjustwidth}
\caption{Results of the hyperparameter search. (\textbf{left}) Variation of BIC with $m$ and $s$ for fixed $\alpha=0.1$. (\textbf{right}) Variation of the BIC with with $m$ and $\alpha$ for fixed $s=1.0$. The white star in each plot indicates the parameters with the lowest BIC across the entire parameter search.\label{fig:hp-results}}
\end{figure}

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=17.0cm]{paper/figures/results/spectral-signatures.png}
\end{adjustwidth}
\caption{Spectral signatures $\psi(\xi_k)$ corresponding to GTM nodes with non-zero responsibility for exemplar spectra corresponding to the rhodamine dye plume (top left), algae (top right), dry grass (bottom left), and open water (bottom right). A pseudo-color image is inset into each plot wit the location of the exemplar spectrum marked with a white circle. \label{fig:spectral-signatures}}
\end{figure}

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=17.0cm]{paper/figures/results/algae.png}
\end{adjustwidth}
\caption{Using the spectral endmember assigned to algae from the trained GTM together with the NS3, we are able to estimate the algal abundance near the shore. (\textbf{left}) NS3 values of pixels below a threshold of 0.4275 corresponding to a total area of $587.3$ $m^2$. (\textbf{right}) A picture of the pond near the shore showing the algae.\label{fig:algae-map}}
\end{figure}

@@

\begin{figure}[t]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=17.0cm]{paper/figures/results/rhodamine.png}
\end{adjustwidth}
\caption{Using the spectral endmember assigned to Rhodamine from the trained GTM together with the NS3, we are able to map the evolution of the dye plume across two UAV flights. (\textbf{left}) The initial dye plume corresponding to a total area of $330.7$ $m^2$. (\textbf{right}) The same dye plume imaged approximately 15 minutes later now subsumes a total area of $1495.6$ $m^2$.\label{fig:rhodamine-map}}
\end{figure}

@@

\noindent 
\begin{tabular}{@{}ll}
UAV & Unmanned Aerial Vehicle \\
GTM & Generative Topographic Mapping  \\
SOM & Self Organizing Map \\ 
HSI & Hyperspectral Image \\ 
PCA & Principal Component Analysis \\ 
tSNE & t-Distributed Stochastic Neighbor Embedding \\
MLJ & Machine Learning in Julia \\ 
VNIR & Visible + Near-Infrared \\
NDWI & Normalized Difference Water Index \\ 
NS3 & Normalized Spectral Similarity Score
\end{tabular}
}
